---
title: "bestNormalize-regression"
author: "Luis Palomero"
date: "`Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bestNormalize-regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}

library(systematicBNR)

```

# Detailed  Linear regression with best-normalize and stepAIC methods to preprocess and select covariables

## Data preprocessing and normalizing

Prior to running regression we must preprocess the data using preprocess_phenotypes and preprocess_effect_variable methods.

```{r load_and_preprocess_data}

set.seed(12345)
n = 500

random_dataset = data.frame(
  sample = sprintf('sample_%s', 1:n),
  signature = (rnorm(n, 10, 3))^2,
  age = floor(runif(n, 45, 70)),
  gender = sample(c('male', 'female'), n, replace = TRUE),
  constant = rep('x', n)
)

random_dataset$signature_derived = log(random_dataset$signature) +  (rnorm(n,0, .25))

summary(random_dataset)

y = systematicBNR::preprocess_effect_variable(random_dataset$signature, allow_orderNorm = FALSE)
plot(
  random_dataset$signature_derived,
  random_dataset$signature
)

phenotypes = systematicBNR::preprocess_phenotypes(
  random_dataset,
  'sample',
  c('age', 'gender', 'constant', 'signature_derived')
)

(method = attr(y, 'chosenTransformation'))

# par(mfrow  = c(2,1))
plot(density(random_dataset$signature), main = 'Original data')
plot(density(y), main = 'Transformed data')
# par(mfrow  = c(1,1))

``` 

Signature value has been normalized using [`r method`]. Transformed data distribucion shape is more "normal" as original one.


## Model fitting

Next step in the pipeline is perform the model fitting using Akaike Incormation Criterion to remove non-informative covariates sistematically.

```{r model_fitting}

original_model = stats::lm(random_dataset$signature ~ ., data = phenotypes)

models = systematicBNR::stepwise_fit_model(y, phenotypes, scope = list(
  lower = ~ 1,
  upper = ~ .
))

summary(original_model)
summary(models$base)
summary(models$fit_stepwise)

```


## Draw model prediction

```{r draw_prediction}


 plot_model <- function(raw_y, phenotypes, predictions, title){

   plot(
     phenotypes$signature_derived,
     raw_y,
     main = title,
     xlab = 'Fake correlated data',
     ylab = 'Random data'
   )
   
   for(i  in 1:length(predictions)){

     aux_df = data.frame(
       x = phenotypes$signature_derived,
       y = predictions[[i]]
     )
     aux_df2 = aux_df[order(aux_df$x),  ]

     lines(
        aux_df2$x,
        aux_df2$y,
        col = i+1
     )
   }
   legend(
      'topleft',
      c('bestNorm/AIC', 'untransformed'),
      pch = 3,
      col = c(2,3),
      cex = 0.8
   )

 }  
 
raw_y = random_dataset$signature
 
predictions = list(
  akaike = attr(models$fit_stepwise, 'untransformed_prediction'),
  raw  = as.numeric(predict(original_model, type = 'response'))
)

plot_model(random_dataset$signature, phenotypes, predictions, 'Comparing base/transformed lm') 


```

# Detailed  Linear regression with best-normalize and stepAIC methods to preprocess and select covariables

Method *systematic_regression* allows run this pipeline sistematically to find candidate covariables.
Below example checks 3 signatures against 'prs' covariable, using 'age-gender-stage-constant' as other phenotype values, generating a list of results combined into a table.

```{r detailed_linerar_regressions}

set.seed(12345)
n = 600

random_dataset = data.frame(
  sample = sprintf('sample_%s', 1:n),
  prs = (rnorm(n, 10, 3))^2,
  age = floor(runif(n, 45, 70)),
  gender = sample(c('male', 'female'), n, replace = TRUE),
  constant = rep('x', n)
)
random_dataset = random_dataset[order(random_dataset$prs),]
random_dataset$stage = c(rep('stage_1', n/3 ), rep('stage_2', n/3 ), rep('stage_3', n/3 ))

signatures = data.frame(
  candidate_a = rnorm(n, 0, 3),
  candidate_b = random_dataset$prs * 2 + rnorm(n, 0, 1),
  candidate_c = runif(n, 10, 20)
)

phenotypes = cbind(random_dataset, signatures)

covariables = c('age', 'gender', 'constant', 'stage')
signature_name = colnames(signatures)[1]

results = lapply(colnames(signatures), function(signature_name, phenotypes){
  systematic_regression(
    phenotypes, 'prs', signature_name, 'sample', covariables
  )  
}, phenotypes)

results = do.call('rbind', results)

print(results)

``` 
